---
title: "Stats 102A - Homework 4 - Output File"
author: "Daren Sathasivam -- UID: 306229580"
output: pdf_document
---

Homework questions and prompts copyright Miles Chen, Do not post, share, or distribute without permission.

# Academic Integrity Statement

By including this statement, I, Daren Sathasivam, declare that all of the work in this assignment is my own original work. At no time did I look at the code of other students nor did I search for code solutions online. I understand that plagiarism on any single part of this assignment will result in a 0 for the entire assignment and that I will be referred to the dean of students.

```{r, error = TRUE}
# source("solutions.R")  # edit with your file name
source("102a_hw_04_script_Daren_Sathasivam.R")  # edit with your file name
print(Player)
print(take_turn)
```

\pagebreak

# Part 1: Test Cases

## do not alter the code for the test cases

### Test Case 1: Space: Go to Jail

```{r, error = TRUE}
dice <- PresetDice$new(
  rolls = c(3,4),
  verbose = TRUE
)
set.seed(16)
player1 <- Player$new(verbose = TRUE, pos = 24)
monopoly <- SpaceTracker$new(verbose = TRUE)

for (i in 1:1) {
  cat("--- Turn", i,"---\n")
  take_turn(player1, monopoly)
  cat("\n")
}

print(setNames(monopoly$counts, 1:40))
```

\pagebreak

### Test Case 2: Chance Card and Doubles Tests: Advance to Go, Reading Railroad, Nearest Railroad, Nearest Utility, No Movement

```{r, error = TRUE}
dice <- PresetDice$new(
  rolls = c(3,4, 4,3, 1,1, 3,4, 5,3),
  verbose = TRUE
)
set.seed(135)
chance <- CardDeck$new(chancedeck, verbose = TRUE)
community <- CardDeck$new(communitydeck, verbose = TRUE)
player1 <- Player$new(verbose = TRUE)
monopoly <- SpaceTracker$new(verbose = TRUE)

for (i in 1:4) {
  cat("--- Turn", i,"---\n")
  take_turn(player1, monopoly)
  cat("\n")
}

print(setNames(monopoly$counts, 1:40))
```
\pagebreak

### Test Case 3: Multiple doubles. Community Chest.
```{r, error = TRUE}
dice <- PresetDice$new(
  rolls = c(3,3, 2,2, 2,1, 3,1), verbose = TRUE)

player1 <- Player$new(verbose = TRUE)
monopoly <- SpaceTracker$new(verbose = TRUE)
for (i in 1:2) {
  cat("--- Turn", i,"---\n")
  take_turn(player1, monopoly)
  cat("\n")
}

print(setNames(monopoly$counts, 1:40))
```

\pagebreak

### Test Case 4: Doubles three times. Three turns in jail.
```{r, error = TRUE}
dice <- PresetDice$new(
  rolls = c(3,3, 3,3, 3,3, 5,6, 5,6, 5,6),
  verbose = TRUE
)

player1 <- Player$new(verbose = TRUE)
monopoly <- SpaceTracker$new(verbose = TRUE)

for (i in 1:4) {
  cat("--- Turn", i,"---\n")
  take_turn(player1, monopoly)
  cat("\n")
}

print(setNames(monopoly$counts, 1:40))
```

\pagebreak

### Test Case 5: After going to Jail, player's turn ends immediately. Rolling doubles while in Jail gets player out of jail.
```{r, error = TRUE}
dice <- PresetDice$new(
  rolls = c(3,3, 1,2, 3,3, 3,4),
  verbose = TRUE
)

player1 <- Player$new(verbose = TRUE, pos = 25)
monopoly <- SpaceTracker$new(verbose = TRUE)

for (i in 1:3) {
  cat("--- Turn", i,"---\n")
  take_turn(player1, monopoly)
  cat("\n")
}

print(setNames(monopoly$counts, 1:40))
```

\pagebreak

### Test Case 6: 20 Predetermined Turns

```{r, error = TRUE}
## You must use these dice for Part 1
dice <- PresetDice$new(
  rolls = c(6,4, 5,3, 3,5, 4,4, 4,4, 2,2, 4,3, 4,4, 1,4,
            3,4, 1,2, 3,6, 5,4, 5,5, 1,2, 5,4, 3,3, 6,1,
            1,1, 2,3, 5,5, 5,4, 4,1, 2,2, 2,4),
  verbose = TRUE
)
set.seed(2)
chance <- CardDeck$new(chancedeck, verbose = TRUE)
community <- CardDeck$new(communitydeck, verbose = TRUE)
# if your chance cards different from mine,
# check to make sure sample(15) returns the following
# > set.seed(2)
# > sample(15)
#  [1]  5  6 14  8  1 11  9  2  3 10  7 12  4 13 15
```


```{r, error = TRUE}
player1 <- Player$new(verbose = TRUE)
monopoly <- SpaceTracker$new(verbose = TRUE)

for (i in 1:20) {
  cat("--- Turn", i,"---\n")
  take_turn(player1, monopoly)
  cat("\n")
}

monopoly$counts
cbind(gameboard, counts = monopoly$counts)
```


# Part 2: 1000 simulated games
```{r, message = FALSE}
library(dplyr)
```


```{r, error = TRUE}
## Use non-verbose random dice for Part 2
set.seed(2)
chance <- CardDeck$new(chancedeck, verbose = FALSE)
community <- CardDeck$new(communitydeck, verbose = FALSE)
dice <- RandomDice$new()

player1 <- Player$new(verbose = FALSE)
player2 <- Player$new(verbose = FALSE)
monopoly <- SpaceTracker$new(verbose = FALSE)

for (g in 1:1000) {
  if (g %% 100 == 0) {
    cat("#### SIMULATING GAME", g, "##### \n")
  }
  for (i in 1:150) {
    take_turn(player1, monopoly)
    take_turn(player2, monopoly)
  }
}

print(setNames(monopoly$counts, 1:40))

results <- cbind(gameboard, tally = monopoly$counts)
results <- cbind(results, rel = monopoly$counts/sum(monopoly$counts))
arrange(results, desc(tally))
print(results)

```


```{r}
# set colors for the bar plot
color_vec <- rep(NA, 40)
color_vec[c(2,4)] <- "chocolate4" # mediterranean, baltic
color_vec[c(7,9,10)] <- "lightskyblue" # oriental, vermont, connecticut 
color_vec[c(12,14,15)] <- "darkorchid2" # st charles, states, virgina
color_vec[c(17,19,20)] <- "darkorange" # st james, tennessee, new york
color_vec[c(22,24,25)] <- "red2" # kentucky, indiana, illinois
color_vec[c(27,28,30)] <- "gold1" # atlantic, ventnor, marvin
color_vec[c(32,33,35)] <- "darkgreen" # pacific, n. carolina, pennsylvania
color_vec[c(38,40)] <- "blue3" # park place, boardwalk
color_vec[c(6,16,26,36)] <- "gray14" # railroads
color_vec[c(13,29)] <- "palegreen1" # utilities
color_vec[c(8,23,37)] <- "salmon2" # chance
color_vec[c(3,18,34)] <- "deepskyblue" # community chest 

barplot(monopoly$counts, 
  main = "Barplot of Frequency of Landing for Each Space",
  xlab = "Space Number", ylab = "Number of Landings",
  las = 2, col = color_vec, names.arg = 1:40, cex.names = 0.65)

# this one sets the y-limits so the max count of jail doesn't shrink everything
barplot(monopoly$counts, 
  main = "Barplot of Frequency of Landing for Each Space",
  xlab = "Space Number", ylab = "Number of Landings",
  las = 2, col = color_vec, names.arg = 1:40, cex.names = 0.65,
  ylim = c(0, sort(monopoly$counts, decreasing = TRUE)[2]))
```

