---
title: "Stats 102A - Homework 2 - Output File"
author: "Daren Sathasivam -- 306229580"
output: pdf_document
---

Homework questions and prompts copyright Miles Chen, Do not post, share, or distribute without permission.

To receive full credit the functions you write must pass all tests. We may conduct further tests that are not included on this page as well.

# Academic Integrity Statement

By including this statement, I, **Daren Sathasivam**, declare that all of the work in this assignment is my own original work. At no time did I look at the code of other students nor did I search for code solutions online. I understand that plagiarism on any single part of this assignment will result in a 0 for the entire assignment and that I will be referred to the dean of students.

```{r, error = TRUE}
source("102a_hw_02_script_Daren_Sathasivam.R")  # edit with your file name
```

\pagebreak

# Part 1: Board representation

Create a single list object called `board` where you store the features of the game board in R.

```{r, error = TRUE}
# chutes, ladders, and board_dim are list objects
board <- list(chutes = chutes <- list(
  start = c(16, 47, 49, 56, 62, 64, 87, 93, 95, 98),
  end = c(6, 26, 11, 53, 19, 60, 24, 73, 75, 78)
), ladders = ladders <- list(
  start = c(1, 4, 9, 21, 28, 36, 51, 71, 80),
  end = c(38, 14, 31, 42, 84, 44, 67, 91, 100)
), dim = board_dim <- list(row = 10, col = 10))
```

\pagebreak

# Part 2: Plot of Game board

```{r, error = TRUE}
# par() should help the plot be more visible. you can adjust this as necessary
board_dim <- list(row = 10, col = 10)
chutes <- list(
  start = c(16, 47, 49, 56, 62, 64, 87, 93, 95, 98),
  end = c(6, 26, 11, 53, 19, 60, 24, 73, 75, 78)
)
ladders <- list(
  start = c(1, 4, 9, 21, 28, 36, 51, 71, 80),
  end = c(38, 14, 31, 42, 84, 44, 67, 91, 100)
)
board <- list(chutes = chutes, ladders = ladders, dim = board_dim)
par(mar = c(0, 0, 0, 0)) 
show_board(board)
```

\pagebreak

# Part 3: Miniboards

Create the `miniboard` objects and plots.

```{r, error = TRUE}
board_dim1 <- list(row = 7, col = 6)
ladders1 <- list(start = c(1, 9, 20, 33), end = c(23, 27, 31, 41))
chutes1 <- list(start = c(16, 30, 35, 40), end = c(6, 18, 24, 26))
miniboard1 <- list(chutes = chutes1, ladders = ladders1, dim = board_dim1)
par(mar = c(1, 7, 1, 7))
show_board(miniboard1)
```

```{r, error = TRUE}
board_dim2 <- list(row = 9, col = 7)
ladders2 <- list(start = c(9, 13, 24, 29, 33, 43), end = c(22, 30, 37, 41, 39, 53))
chutes2 <- list(start = c(16, 31, 35, 62), end = c(3, 15, 21, 48))
miniboard2 <- list(chutes = chutes2, ladders = ladders2, dim = board_dim2) 
par(mar = c(1, 8, 1, 8))
show_board(miniboard2)
```

```{r, error = TRUE}
board_dim3 <- list(row = 9, col = 7)
ladders3 <- list()
chutes3 <- list()
miniboard3 <- list(chutes = chutes3, ladders = ladders3, dim = board_dim3)
par(mar = c(1, 8, 2, 8))
show_board(miniboard3)
```

\pagebreak

# Part 4: Verbose output of one single player game

```{r, error = TRUE}
set.seed(5)
play_solo(board, verbose = TRUE)
```

\pagebreak

# Part 5: Monte Carlo Simulation Study

```{r, error = TRUE}
# run 10,000 games
simulation <- replicate(10000, play_solo(board), simplify = FALSE)
```

-   Create a histogram (breaks = 50) of the turns.

```{r}
total_turns <- sapply(simulation, function(x){
  x$turns
})
hist(total_turns, breaks = 50, main = "Histogram of turns to complete a game of Chutes and Ladders", xlab = "Turns")
```

-   Find the minimum number of turns. How many times out of 10,000 did a game finish with the minimum number of turns?

```{r}
min_turns <- min(total_turns)
# Minimum number of turns in a game
min_turns
# For amount of games that finished with the minimum amount of turns
sum(total_turns == min_turns)
```

-   Find the maximum number of turns.

```{r}
max_turns <- max(total_turns)
max_turns
```

-   What is the median number of turns?

```{r}
median_turns <- median(total_turns)
median_turns
```

-   What is the mean number of turns?

```{r}
mean_turns <- mean(total_turns)
mean_turns
```

-   What proportion of games take 100 or more turns to complete?

```{r}
prop_morethan_100 <- mean(total_turns >= 100)
prop_morethan_100
```

-   What proportion of games take 10 or fewer turns to complete?

```{r}
prop_lessthan_10 <- mean(total_turns <= 10)
prop_lessthan_10
```

-   What proportion of games utilize ladder 9 (the shortcut to win on space 80)?

```{r}
chute_tallies <- tapply(unlist(lapply(simulation, function (x) {
  x$chute_tally
})), rep(seq_along(board$chutes$start), times = 10000), sum)
ladder_tallies <- tapply(unlist(lapply(simulation, function (x) {
  x$ladder_tally
})), rep(seq_along(board$ladders$start), times = 10000), sum)

prop_ladder9 <- ladder_tallies[9] / 10000
prop_ladder9
```

-   Barplot of the relative frequency of how often each chute is utilized.
```{r}
barplot(chute_tallies / 10000, main = "Relative Frequency of Chute Usage", xlab = "Chute Number", ylab = "Frequency", col = "red")
```

-   Barplot of the relative frequency of how often each ladder is utilized.
```{r}
barplot(ladder_tallies / 10000, main = "Relative Frequncy of Ladder Usage", xlab = "Ladder Number", ylab = "Frequency", col = "green")
```

\pagebreak

# Extra Credit Challenge
Pages to keep note of: \
* https://r4ds.had.co.nz/data-visualisation.html \
* https://r-charts.com/correlation/heat-map-ggplot2/ \
* https://datacarpentry.org/R-ecology-lesson/04-visualization-ggplot2.html \
* https://stats.stackexchange.com/questions/26722/calculate-transition-matrix-markov-in-r \
* https://www.datacamp.com/tutorial/markov-chain-analysis-r \
* https://ggplot2.tidyverse.org/reference/aes.html \
```{r}
library(ggplot2)

roll1 <- matrix(0, nrow = 100, ncol = 100)
# Find chute or ladder end --> find_special from part 5
special_end <- function(pos, board) {
  if (pos %in% board$chutes$start) {
    return(board$chutes$end[which(board$chutes$start == pos)])
  } else if (pos %in% board$ladders$start) {
      return(board$ladders$end[which(board$ladders$start == pos)])
  } else {
    return(pos)
  }
}
# Fill in the matrix
for (start in 1:100) {
  for (roll in 1:6) {
    end <- min(start + roll, 100)
    end <- special_end(end, board)
    roll1[start, end] <- roll1[start, end] + (1 / 6)
  }
}
# Calculate roll 2 and 3
roll2 <- roll1 %*% roll1
roll3 <- roll2 %*% roll1
# Create the heat map
plot_heat_map <- function(transition_mat, title) {
  transition_df <- as.data.frame(as.table(transition_mat))
  # Plot df
  ggplot(transition_df, aes(Var1, Var2, fill = Freq)) +
    geom_tile() + 
    scale_fill_gradient(low = "yellow", high = "red") +
    labs(title = title, x = "Start Position", y = "End Position")
}
# Plot heat maps on rolls
par(mfrow = c(1, 3))
plot_heat_map(roll1, "Heat Map After 1 Roll")
plot_heat_map(roll2, "Heat Map After 2 Rolls")
plot_heat_map(roll3, "Heat Map After 3 Rolls")

```

