---
title: "102A_Disc_W3-6"
author: "Daren Sathasivam"
date: "2024-01-23"
output: html_document
---

# Week 3 Discussion
## HW 2
### Part 1
```{r}
board <- list(dimx, dimy, chutes, ladders)
rbind(c(y, 7)) # ex
rbind(c(y1, y2))
```

### Part 2: Board visualization
1. Segments (x0, y0, x1, y1) - draw a segment from the points
2. text (x, y, str) - creates a text box with given string
3. arrows (x0, y0, x1, y1) - draws an arrows from given points
**STEPS**
1st step: Contruct the grid using segments \
* dimx + 1 --> vertical lines
* dimy + 1 --> horizontal lines
2nd step: Fill the grid with numbers \
* find_coord(x, dimx, dimy)
* invert on even rows: \
0 <--> (0.5, 1.5)\
10 <--> (4.5, 1.5)\
* check if the row is even \
(x, y) --> (dimx - x, y)\

### Part 3: Test Miniboards
Go through variations of dimensions and chutes and ladders

### Part 4: Game Simulation
```{r}
play_solo(board, roll)
```
Exit condition: pos = 100\
each turn spin = 6d\
  check edge case pos > 100\
  pos = pos + spin\
  check if on chute/ladder square\
    * use match(pos, first column of ladders/chutes)
    * Record number of turns and chutes and ladders
### Part 5: MC Simulations
num of turns\
chute count\
ladder count\
move logs\
10000 --> set random seed \
each time you start the simulations \
\# turns_total <- c(...)\
simulation(max, median) -- summary(...)

# Week 4 Discussion
## HW 2
### Part 1
```{r}
(unique(c(1, 1, 1, 2, 3, 4, 7)))
```
* Get familiar with deplyr

### Part 2 more deplyr
* everything is formatted as date
* find users that visited more than once
  * use mutate() then rowwise() to calculate number of visits
  * filter() to find ids with more than one visit
* write helper function
  * total_days_between_visits(visit1, visit2, visit3, visit4, visit5)
* get total number of visits and get total difference in days

### Part 3 rvest
* webscraping
* extract hyperlinks in active franchise's
  * vectorize
* scrape franchise history table
* get selector gadget
  * use selector gadget to find CSS heading
  * click to red to deselct node
  * once selected, copy the node from bar above -- >html_nodes("selector_gadget_node")
  * html_attr("href") -- gets hyperlinks
  * html_text() -- to get text
* loop through all href and visit each site and gather data
  * correct tag is #franchise_years
* merge tables with all teams
* check dim (2804x22)

### Part 4 dplyr to summarize data
* filter data
* summary statistics


### Part 5: Regex
* letters / digits
* \\w for alphabetical
* \\d for digits
* [:alpha:] - [a-zA-Z]
* [:alnum:] - any alphanumerical class
* *(wildcard) - represents any non-whitespace character
* character ranges [a-zA-Z]
* repetitions - [aaabc]
* Quanitities - 
  
