---
title: "102ADiscNotesW2-4"
author: "Daren Sathasivam"
date: "2024-01-16"
output: html_document
---

# Week 2
## January 16th, 2024(Tues)

### Code Chunks
* eval = FALSE - prevents code from being evaluated
* include = FALSE - runs the code but doesn't show the code or results in the final document
* echo = FALSE - prevents code from showing in the final output, but does not show the results of the code
* message = FALSE - prevents messages or warnings from appearing in the final document

Example of sourcing an R script
```{r}
# source("demo.R")
# prodratio(5, 3)
```

### Including Plots

### Try-Catch for Part 1 HW1
```{r}
my_log <- function(x) {
  tryCatch( {
    result = log(x)
    return(result)
  }
  error = function(e) {
    message("An Error Occurred!")
    print(e)
  }
  warning=function(w){
    # message("A warning occurred")
    # print(w)
    return(0)
  }
  )
}
```


### Part 2 HW1
```{r}
# x = 180
# d = 2
# --> res = 2
# x = 90
# d = 2
# --> res = 2, 2
# x = 45
# d = 2
# 45 / 2 != N --> increment d
# x = 45
# d = 3
# res = 2, 2, 3
# x = 15
# d = 3
# res = 2, 2, 3, 3
# x = 5
# d = 3 + 1 + 1
# res = 2, 2, 3, 3, 5
```

### Part 3 HW1
```{r}

```

